Understanding LIME

### 2019-08-24 README.Rmd
worked on organizing journals

- Paper
    - I found ridge regression models in Python code (see journal 04)
    - Show the plots using the ridge regression in journal 09
    - What should I do about computing the generalized R2 value?
    - Show this paper: https://arxiv.org/pdf/1806.07498.pdf
    - Can you help to show me your vision with the minipage format?
    - I could not find a locality metric output from Python. There was a section in their vignettes that seemed to be referring to some sort of assessment about the models, but I'm not sure what was happening. https://marcotcr.github.io/lime/tutorials/Tutorial%20-%20continuous%20and%20categorical%20features.html
- Graphics Group
    - email list (should I be added as an admin?)
    - send out another email and plan for first week?
    - updates on schedule since last meeting

@@

- keep the LIME procedure at a higher level in the paper
- change to computing the $\mathcal{L}$ metric from the paper
- add figure descriptions in both caption and text
- go back through LIME procedure and update with new understanding (such as d' can be larger than d)
- add an assert_that statement in the code to check if the LIME package is changing
- read through papers giving an overview of the current explanation methods

>>>>

- Paper
    - Can you re-explain your vision with the minipage format?
- Internship
    - Dixon said he is fine with me switching to 10 hours in the spring (as long as Dr. Caragea agrees)
- Graphics Group
    - email list (should I be added as an admin?)
    - emailed Min (waiting for a response)
    - show updated schedule
    - Susan's idea (could we ask Joe, Erin, and/or Charlotte if they would be interested in preparing this talk?)

@@

- finish organizing journals and update contents in readme
- send out an email about graphics group plan for the first week
- finish reading new paper
- set up meeting with Danny on Friday once Heike sends me her availability
- add generalized R2 values to plots
- code up Laugel et al's work
- point out differences and similarities from the other paper (such as normal sampling)
- keep the LIME procedure at a higher level in the paper
- change to computing the $\mathcal{L}$ metric from the paper
- add figure descriptions in both caption and text
- go back through LIME procedure and update with new understanding (such as d' can be larger than d)
- add an assert_that statement in the code to check if the LIME package is changing
- read through papers giving an overview of the current explanation methods

### 2019-08-24 journals/00-objectives_and_ideas/00-objectives_and_ideas.Rmd
worked on organizing journals

---
title: "00-objectives_and_ideas"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

# Overview

This journal explains the objectives of the research project on applying LIME to the Hamby bullet data that is detailed in this series of journals.

# Objectives of the Research Project

It is the hope that this research project will result in two papers. One will focus on a statistical assessment of LIME and the other will focus on the explanation of how the application of LIME to the Hamby bullet data can help firearm examiners understand the random forest model predictions.

### Statistical Paper Presenting Visual Diagnostic Tools for LIME

Write a paper that introduces some visual diagnostics for LIME to assess its performance. Key points that will be made:

1. explainer model is really bad (probably due to the binning)
2. local explanations at not local

### Statistical Paper Assessing LIME

While applying LIME to the bullet matching dataset, we found that the explanations were not as good as we wanted them to be. This has led us to try to understand why the LIME algorithm is not working well and to propose some new methods to improve the algorithm. This paper will discuss isusses we found with the current LIME agorithm, describe our proposed approaches, and compare the results.

### Firearm Examiner's Paper

@@

# Concerns with LIME

@@

# General Thoughts about LIME

>>>>

---
title: "Objectives and Ideas"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

This journal explains the objectives and ideas for the research project on applying LIME to the Hamby bullet data that is detailed in this series of journals.

# Objectives

### Overall

- Understand how LIME works
- Consider ways to improvement LIME
- Apply LIME (or some improved version of LIME) to interpret the random forest model fit to the Hamby bullet data

### Possible Papers

Here are ideas for possible papers.

**Statistical Paper Presenting Visual Diagnostic Tools for LIME**

This paper would assess the explainations produced by LIME. It would introduce some visual diagnostics for LIME to assess its performance. Key points that will be made:

1. Explainer model is really bad (probably due to the binning)
2. Local explanations at not local

**Statistical Paper Assessing LIME**

This paper would be an explanation of how the application of LIME to the Hamby bullet data can help firearm examiners understand the random forest model predictions. When we first applied LIME to the bullet matching dataset, we found that the explanations were not as good as we wanted them to be. This has led us to try to understand why the LIME algorithm is not working well. We even tried some new sampling methods to improve the algorithm. They mde a bit of improvement, but the improvements could be much better. This paper could discuss isusses we found with the current LIME agorithm, describe our proposed approaches, and compare the results.

**Firearm Examiner's Paper**

@@

# Concerns

@@

# General Thoughts

### 2019-08-24 journals/01-hamby_data_and_rtrees/01-hamby_data_and_rtrees.Rmd
worked on organizing journals

---
title: "01-training and testing data"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

@@

# Overview

This journal includes a description of the Hamby study datasets, and the code that is used to clean the Hamby training and testing datasets.

```{r}
# Load libraries
library(tidyverse)
library(plotly)
library(randomForest)
```

Both the training and the testing data will require the use of a vector of the names of the features used in the `rtrees` random forest included in the bulletr library. I create such a vector below to use in this journal.

```{r}
# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(bulletr::rtrees$importance)
```

# Description of the Hamby Data

@@

### Variable Definitions

The definitions of the variables...

@@

# Training Data

@@

### The Raw Data

@@

### Considering the Number of Rows in the Raw Data

@@

### Understanding the Missing Observations

@@

### Cleaning the Trainng Data

@@

### Removing Bullets with Tank Rash

@@

### Saving the Training Data

@@

# Testing Data

### The Raw Data

@@

### Cleaning the Testing Data

@@

### Saving the Testing Data

@@

>>>>

---
title: "Hamby Data and `rtrees`"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

@@

This journal includes a description of the Hamby datasets and the random forest model `rtrees` that will be used in this research project. Additionally, the specification and cleaning of the training and testing datasets are documented, and some initial visualizations of the random forest model are included.

```{r}
# Load libraries
library(bulletr)
library(cowplot)
library(plotly)
library(randomForest)
library(tidyverse)
```

# Hamby Data

### Description

@@

Both the training and the testing data will require the use of a vector of the names of the features used in the `rtrees` random forest included in the bulletr library. I created such a vector to use in this journal, and this vector is printed below. It includes nine features.

```{r}
# Obtain features used when fitting the rtrees random forest
rf_features <- rownames(bulletr::rtrees$importance)
rf_features
```

At some point, I would like to write down definitions of all the features.

@@

### Training Data

@@

**Raw Data**

@@

**Number of Rows**

@@

**Missing Observations**

@@

**Data Cleaning**

@@

**Removing Bullets with Tank Rash**

@@

**Saving the Data**

@@

### Testing Data

**Raw Data**

@@

**Data Cleaning**

@@

**Saving the Data**

@@

### Plots of Features

##### Distributions by `samesource`

I created the following two plots of the training data as suggested by Heike. The histograms below show the distributions of the features used in the random forest `rtrees`. The histograms are filled by the `samesource` variable, which is the truth of whether or not the comparison is from the same barrel and land. The default histograms make it hard to compare the distributions of the matches and non-matches since there are many more comparisons that have `samesource == FALSE`.

```{r}
# Create plots of the feature distributions colored by samesource for the training data
hamby173and252_train %>% 
  select(rf_features, samesource) %>%
  gather(key = feature, value = value, 1:9) %>%
  select(feature, value, samesource) %>%
  ggplot(aes(x = value, fill = samesource)) + 
  geom_histogram(bins = 30) + 
  facet_wrap( ~ feature, scales = "free") +
  labs(x = "Variable Value", y = "Frequency", fill = "Same Source?",
       title = "Hamby 173 and 252 Training Data") +
  theme_bw() +
  theme(legend.position = "bottom")
```

By setting `position = "fill"` in the `geom_histogram` function, it is easier to compare the matches and non-matches. <span style="color:teal"> These plots could be used in the future to hand select the bins for lime. Additionally, fitting a logistic regression to this data could also be used to determine the LC50, LC10, ad LC90, which could be used as the bins for lime. </span>

```{r}
# Create plots of the feature distributions colored by samesource for the training data
# using the position = "fill" option
hamby173and252_train %>% 
  select(rf_features, samesource) %>%
  gather(key = feature, value = value, 1:9) %>%
  select(feature, value, samesource) %>%
  ggplot(aes(x = value, fill = samesource)) + 
  geom_histogram(position = "fill", bins = 30) + 
  facet_wrap( ~ feature, scales = "free") +
  labs(x = "Variable Value", y = "Proportion", fill = "Same Source?",
       title = "Hamby 173 and 252 Training Data") +
  theme_bw() +
  theme(legend.position = "bottom")
```

The plots below have the same structure, but they are created with the Hamby 224 testing data. I chose to not separate the testing data by sets, but this is something that could be done later if necessary.

```{r}
# Create plots of the feature distributions colored by samesource for the testing data
hamby224_test %>% 
  select(rf_features, samesource) %>%
  gather(key = feature, value = value, 1:9) %>%
  select(feature, value, samesource) %>%
  ggplot(aes(x = value, fill = samesource)) + 
  geom_histogram(bins = 15) + 
  facet_wrap( ~ feature, scales = "free") +
  labs(x = "Variable Value", y = "Frequency", fill = "Same Source?",
       title = "Hamby 224 Testing Data") +
  theme_bw() +
  theme(legend.position = "bottom")
```

```{r}
# Create plots of the feature distributions colored by samesource for the testing 
# data using the position = "fill" option
hamby224_test %>% 
  select(rf_features, samesource) %>%
  gather(key = feature, value = value, 1:9) %>%
  select(feature, value, samesource) %>%
  ggplot(aes(x = value, fill = samesource)) + 
  geom_histogram(position = "fill", bins = 15) + 
  facet_wrap( ~ feature, scales = "free") +
  labs(x = "Variable Value", y = "Proportion", fill = "Same Source?",
       title = "Hamby 224 Testing Data") +
  theme_bw() +
  theme(legend.position = "bottom")
```

##### Correlations

I made these plots to look at the correlation between features in the training data within the TRUE and FALSE cases of samesource. The features are highly correlated for the match comparisons. It is clear that the variables are more correlated with the match comparisons than the non-match comparisons. However, there are still some variables that are relatively highly correlation with the non-match comparisons.

```{r fig.height = 4, fig.width = 9}
# Create a correlation heatmap of the match comparisons in the training data
cor_match <- hamby173and252_train %>%
  select(rf_features, samesource) %>%
  filter(samesource == TRUE) %>%
  select(-samesource) %>%
  cor() %>%
  reshape2::melt() %>%
  mutate(Var1 = factor(Var1, levels = c("ccf", "cms", "matches", "rough_cor", "sum_peaks",
                                        "D", "sd_D", "mismatches", "non_cms")),
         Var2 = factor(Var2, levels = c("ccf", "cms", "matches", "rough_cor", "sum_peaks",
                                        "D", "sd_D", "mismatches", "non_cms"))) %>%
  ggplot(aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient2(limits = c(-1, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") + 
  labs(x = "", y = "", fill = "Correlation",
       title = "Match Comparisons")

# Create a correlation heatmap of the non-match comparisons in the training data
cor_nonmatch <- hamby173and252_train %>%
  select(rf_features, samesource) %>%
  filter(samesource == FALSE) %>%
  select(-samesource) %>%
  cor() %>%
  reshape2::melt() %>%
  mutate(Var1 = factor(Var1, levels = c("ccf", "cms", "matches", "rough_cor", "sum_peaks",
                                        "D", "sd_D", "mismatches", "non_cms")),
         Var2 = factor(Var2, levels = c("ccf", "cms", "matches", "rough_cor", "sum_peaks",
                                        "D", "sd_D", "mismatches", "non_cms"))) %>%
  ggplot(aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient2(limits = c(-1, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(x = "", y = "", fill = "Correlation",
       title = "Non-Match Comparisons")

# Create a title for the panel of plots
cor_title <- ggdraw() +
  draw_label("Correlation of Feature Variables in the Training Data", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             hjust = 0.5)

# Create a joined legend for the panel of plots
cor_legend <- get_legend(cor_match + theme(legend.position = "right"))

# Create the panel of plots
plot_grid(cor_title, 
          plot_grid(cor_match, cor_nonmatch, cor_legend, ncol = 3, rel_widths = c(2, 2, 0.5)),
          ncol = 1,
          rel_heights = c(0.25, 3))
```

The plots below show the correlations for the testing data. The patterns in the plots look really similar to ones of the training data.

```{r fig.height = 4, fig.width = 9}
# Create a correlation heatmap of the match comparisons in the training data
cor_match_test <- hamby224_test %>%
  select(rf_features, samesource) %>%
  filter(samesource == TRUE) %>%
  select(-samesource) %>%
  cor() %>%
  reshape2::melt() %>%
  mutate(Var1 = factor(Var1, levels = c("ccf", "cms", "matches", "rough_cor", "sum_peaks",
                                        "D", "sd_D", "mismatches", "non_cms")),
         Var2 = factor(Var2, levels = c("ccf", "cms", "matches", "rough_cor", "sum_peaks",
                                        "D", "sd_D", "mismatches", "non_cms"))) %>%
  ggplot(aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient2(limits = c(-1, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") + 
  labs(x = "", y = "", fill = "Correlation",
       title = "Match Comparisons")

# Create a correlation heatmap of the non-match comparisons in the training data
cor_nonmatch_test <- hamby224_test %>%
  select(rf_features, samesource) %>%
  filter(samesource == FALSE) %>%
  select(-samesource) %>%
  cor() %>%
  reshape2::melt() %>%
  mutate(Var1 = factor(Var1, levels = c("ccf", "cms", "matches", "rough_cor", "sum_peaks",
                                        "D", "sd_D", "mismatches", "non_cms")),
         Var2 = factor(Var2, levels = c("ccf", "cms", "matches", "rough_cor", "sum_peaks",
                                        "D", "sd_D", "mismatches", "non_cms"))) %>%
  ggplot(aes(x = Var1, y = Var2, fill = value)) + 
  geom_tile() + 
  scale_fill_gradient2(limits = c(-1, 1)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "none") +
  labs(x = "", y = "", fill = "Correlation",
       title = "Non-Match Comparisons")

# Create a title for the panel of plots
cor_title_test <- ggdraw() +
  draw_label("Correlation of Feature Variables in the Testing Data", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             hjust = 0.5)

# Create a joined legend for the panel of plots
cor_legend_test <- get_legend(cor_match_test + theme(legend.position = "right"))

# Create the panel of plots
plot_grid(cor_title_test, 
          plot_grid(cor_match_test, cor_nonmatch_test, cor_legend_test, 
                    ncol = 3, rel_widths = c(2, 2, 0.5)),
          ncol = 1,
          rel_heights = c(0.25, 3))
```

# Random Forest: `rtrees`

### Description

The random forest model that was fit to the Hamby training dataset is discussed in the paper "Automatic Matching of Bullet Land Impressions" by E. Hare Et. Al. The model is available through the GitHub version of the `bulletr` package. The link to the GitHub repository for `bulletr` is https://github.com/CSAFE-ISU/bulletr.

The random forest model is stored in the package `bulletr` as a data object called `rtrees`. The paper does not include many details about the model fitting process, so I looked at some of the model features, which are printed below. These values told me that the model was fit using the predictive features of `ccf`, `rough_cor`, `D`, `sd_D`, `matches`, `mismatches`, `cms`, `non_cms`, and `sum_peaks` with 300 trees and the parameter of `mtry` set to 3, which is the "number of variables randomly sampled as candidates at each split" as explained in the `randomforest` package documentation.

```{r}
# Importance values from the random forest
rtrees$importance

# The number of trees used in the random forest
rtrees$ntree

# "Number of variables randomly sampled as candidates at each split"
rtrees$mtry
```

### Visualizations

##### Random Forest Scores

**Facet Version**

The plot below is the model for the one that will be used in the app for exploring the `lime` explanations from the bullet matching data. This is the data from set 1 of the testing dataset.

```{r}
# Heatmap of rfscore for each comparison in set 1
hamby224_test %>%
  filter(set == "Set 1") %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2,
             text = paste('Bullets Compared: ', bullet1, "-", land1, 
                          "vs", bullet2, "-", land2,
                          '\nRandom Forest Score: ', 
                          ifelse(is.na(rfscore), "Missing due to tank rash", rfscore)))) +
  geom_tile(aes(fill = rfscore)) +
  facet_grid(bullet2 ~ bullet1, scales = "free") +
  theme_minimal() +
  scale_fill_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank()) +
  labs(x = "", y = "", fill = "RF Score")
```

**Individual Plots Version**

I also wrote some code to create the above plot using individual plots instead of facets. This is included below. (It is a bit outdated compared to my newer version of the plot.)

```{r}
# Code for creating the heatmap plots using individual plots
p1 <- hamby224_test %>%
  filter(set == "Set 1", bullet1 == "Known 1", bullet2 == "Known 1") %>%
  select(case, bullet1, bullet2, land1, land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2)) +
  geom_tile(aes(fill = rfscore)) +
  theme_minimal() +
  scale_fill_gradient2(low = "grey", high = "orange", midpoint = 0.5) +
  labs(x = "Land 1", y = "Land 2", fill = "RF Score") + 
  theme(legend.position = "none")

p2 <- hamby224_test %>%
  filter(set == "Set 1", bullet1 == "Known 1", bullet2 == "Known 2") %>%
  select(case, bullet1, bullet2, land1, land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2)) +
  geom_tile(aes(fill = rfscore)) +
  theme_minimal() +
  scale_fill_gradient2(low = "grey", high = "orange", midpoint = 0.5) +
  labs(x = "Land 1", y = "Land 2", fill = "RF Score") + 
  theme(legend.position = "none")

p3 <- hamby224_test %>%
  filter(set == "Set 1", bullet1 == "Known 1", bullet2 == "Questioned") %>%
  select(case, bullet1, bullet2, land1, land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2)) +
  geom_tile(aes(fill = rfscore)) +
  #facet_grid(bullet1 ~ bullet2, scales = "free") +
  theme_minimal() +
  scale_fill_gradient2(low = "grey", high = "orange", midpoint = 0.5) +
  labs(x = "Land 1", y = "Land 2", fill = "RF Score") + 
  theme(legend.position = "none")

p4 <- ggplot() + geom_blank() + theme_classic()

p5 <- hamby224_test %>%
  filter(set == "Set 1", bullet1 == "Known 2", bullet2 == "Known 2") %>%
  select(case, bullet1, bullet2, land1, land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2)) +
  geom_tile(aes(fill = rfscore)) +
  #facet_grid(bullet1 ~ bullet2, scales = "free") +
  theme_minimal() +
  scale_fill_gradient2(low = "grey", high = "orange", midpoint = 0.5) +
  labs(x = "Land 1", y = "Land 2", fill = "RF Score") + 
  theme(legend.position = "none")

p6 <- hamby224_test %>%
  filter(set == "Set 1", bullet1 == "Known 2", bullet2 == "Questioned") %>%
  select(case, bullet1, bullet2, land1, land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2)) +
  geom_tile(aes(fill = rfscore)) +
  #facet_grid(bullet1 ~ bullet2, scales = "free") +
  theme_minimal() +
  scale_fill_gradient2(low = "grey", high = "orange", midpoint = 0.5) +
  labs(x = "Land 1", y = "Land 2", fill = "RF Score") + 
  theme(legend.position = "none")

p7 <- ggplot() + geom_blank() + theme_classic()

p8 <- ggplot() + geom_blank() + theme_classic()

p9 <- hamby224_test %>%
  filter(set == "Set 1", bullet1 == "Questioned", bullet2 == "Questioned") %>%
  select(case, bullet1, bullet2, land1, land2, rfscore) %>%
  distinct() %>%
  ggplot(aes(x = land1, y = land2, label = bullet1, label2 = bullet2)) +
  geom_tile(aes(fill = rfscore)) +
  #facet_grid(bullet1 ~ bullet2, scales = "free") +
  theme_minimal() +
  scale_fill_gradient2(low = "grey", high = "orange", midpoint = 0.5, limits = c(0,1)) +
  labs(x = "Land 1", y = "Land 2", fill = "RF Score")

style(subplot(p1, p2, p3, p4, p5, p6, p7, p8, p9, 
              nrows = 3, 
              titleX = TRUE, 
              titleY = TRUE, 
              margin = 0.03),
      hoverinfo = "skip",
      traces = 7)
```

##### Parallel Coordinate Plots

```{r}
# Create a dataframe with the random forest features ordered by importance
rf_features_ordered <- data.frame(feature = rownames(bulletr::rtrees$importance), 
                                  MeanDecreaseGini = bulletr::rtrees$importance) %>%
  arrange(desc(MeanDecreaseGini))
```

Below is a parallel coordinate plot of the training dataset. It is facetted by `samesource` and colored by `rfscore`. The features are sorted by the variable importance scores (descending from left to right) from the `rtrees` model. It is interesting to see that the matches with low rfscores have feature values that are similar to the nonmatches.

```{r fig.width = 10}
hamby173and252_train %>%
  mutate(case = 1:n()) %>%
  select(case, rfscore, samesource, ccf:sum_peaks) %>%
  mutate_at(rf_features, scale) %>%
  gather(variable, value, 4:12) %>%
  mutate(variable = factor(variable, levels = rf_features_ordered$feature)) %>%
  ggplot(aes(x = variable, y = value, group = case, color = rfscore)) + 
  geom_point() + 
  geom_line(alpha = 0.2) + 
  facet_grid( ~ samesource) + 
  scale_color_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Feature (sorted by rtrees variable importance)", 
       y = "Standardized Feature Values")
```

Below is a parallel coordinate plot of the testing dataset. It is facetted by `samesource` and colored by `rfscore`. You can see a clear distinction between the trends of the feature values from the nonmatches and matches. It is interesting to consider the cases in set 1 that are matches but were predicted to be nonmatches by the model. These are the cases colored in grey, and their feature values are very different from the observations that were predicted to be matches by the model. The nonmatch predictions have values that are closer to the nonmatches.

```{r fig.width = 10}
hamby224_test %>%
  na.omit() %>%
  select(case, set, rfscore, samesource, ccf:sum_peaks) %>%
  mutate_at(rf_features, scale) %>%
  gather(variable, value, 5:13) %>%
  mutate(variable = factor(variable, levels = rf_features_ordered$feature)) %>%
  ggplot(aes(x = variable, y = value, group = case, color = rfscore)) + 
  geom_point() + 
  geom_line(alpha = 0.5) + 
  facet_grid(set ~ samesource) + 
  scale_color_gradient2(low = "darkgrey", high = "darkorange", midpoint = 0.5) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) + 
  labs(x = "Feature (sorted by rtrees variable importance)", 
       y = "Standardized Feature Values")
```

##### Biplot

```{r eval = FALSE}
pca_set1 <- princomp(x = hamby224_test %>% 
                       filter(set == "Set 1") %>% 
                       select(rf_features) %>% 
                       na.omit(), 
                     cor = TRUE)
biplot(pca, col = hamby224_test %>% 
                       filter(set == "Set 1") %>% 
                       pull(rfscore))
```

### 2019-08-24 journals/02-understanding_lime/02-understanding_lime.Rmd
worked on organizing journals

---
title: "Lime Algorithm and Proposed Methods"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE)
```

# Overview

@@

```{r}
# Source functions
source("../../code/helper_functions.R")
```

@@

>>>>

---
title: "Understanding LIME"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, eval = TRUE, cache = TRUE)
```

@@

```{r}
# Load packages
library(cowplot)
library(furrr)
library(future)
library(glmnet)
library(gower)
library(lime)
library(randomForest)
library(tidyverse)

# Source functions
source("../../code/helper_functions.R")
```

@@

# Concept Visualizations

The images in this section were created using simulated data that I knew had a local trend around the point of interest. I consider both a linear regression and a ridge regression as the explainer model.

### Data and Model

This code generates the data and predictions from a hypothetical complex model.

```{r}
# Generate features and predictions from a hypothetical complex model
set.seed(20190624)
lime_data <- data.frame(feature1 = sort(runif(250, 0, 1)),
                        feature2 = sample(x = 1:250, 
                                          size = 250, 
                                          replace = FALSE)) %>%
  mutate(prediction = if_else(feature1 >= 0 & feature1 < 0.1, 
                              (0.3 * feature1) + rnorm(n(), 0, 0.01), 
                      if_else(feature1 >= 0.1 & feature1 < 0.3, 
                              rbeta(n(), 1, 0.5), 
                      if_else(feature1 >= 0.3 & feature1 < 0.5, 
                              sin(pi* feature1) + rnorm(n(), 0, 0.5),
                      if_else(feature1 >= 0.5 & feature1 < 0.8, 
                              -(sin(pi* feature1) + rnorm(n(), 0, 0.1)) + 1,
                      if_else(feature1 >= 0.8 & feature1 < 0.9, 
                              0.5 + runif(n(), -0.5, 0.5), 
                              0.5 + rnorm(n(), 0, 0.3)))))))
```

```{r}
# Specify a prediction of interest
prediction_of_interest <- data.frame(feature1 = 0.07, 
                                     feature2 = 200,
                                     prediction = 0.05, 
                                     color = factor("Prediction of Interest"))
```

```{r}
# Compute the distance between the prediction of interest and 
# all other observations
lime_data$distance <- (1 - gower_dist(x = prediction_of_interest, 
                                      y = lime_data))^20
```

### Linear Regression Explainer

These visualizations are based on a weighted linear regression model used as the explainer.

```{r}
# Fit the interpretable explainer model (weighted linear regression)
linear_explainer <- lm(formula = prediction ~ feature1 + feature2, 
                       data = lime_data, 
                       weights = distance)
```

```{r}
# Extract the coefficients from the explainer
linear_b0 <- coef(linear_explainer)[[1]]
linear_b1 <- coef(linear_explainer)[[2]]
linear_b2 <- coef(linear_explainer)[[3]]
```

```{r}
# Predictions versus feature 1
linear_plot_feature1 <- ggplot(data = lime_data,
                        mapping = aes(x = feature1, 
                                      y = prediction, 
                                      size = distance)) + 
  geom_point(color = "grey30",
             alpha = 0.75) + 
  geom_abline(intercept = linear_b0 +
                linear_b2*prediction_of_interest$feature2, 
              slope = linear_b1, 
              size = 1) +
  geom_point(data = prediction_of_interest,
             mapping = aes(x = feature1, y = prediction),
             fill = "#EFB084",
             color = "black",
             size = 5, 
             shape = 23) +
  theme_linedraw(base_family = "Times") +
  labs(x = "Feature 1", 
       y = "Black-Box Prediction") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none")
```

```{r}
# Predictions versus feature 1
linear_plot_feature2 <- ggplot(data = lime_data,
                        mapping = aes(x = feature2, 
                                      y = prediction, 
                                      size = distance)) + 
  geom_point(color = "grey30",
             alpha = 0.75) + 
  geom_abline(intercept = linear_b0 + 
                linear_b1*prediction_of_interest$feature1,
              slope = linear_b2,
              size = 1) +
  geom_point(data = prediction_of_interest,
             mapping = aes(x = feature2, y = prediction),
             fill = "#EFB084",
             color = "black",
             size = 5, 
             shape = 23) +
  theme_linedraw(base_family = "Times") +
  labs(x = "Feature 2", 
       y = "Black-Box Prediction") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none")
```

```{r}
# Plot that creates a size legend
size_legend_plot <- lime_data %>%
  gather(key = feature, value = value, 1:2) %>%
  ggplot(mapping = aes(x = value, 
                       y = prediction,
                       size = distance)) + 
  geom_point(color = "grey30",
             alpha = 0.75) +
  labs(size = "Weight") +
  theme_linedraw(base_family = "Times") +
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank())

# Extract the legend
size_legend <- get_legend(size_legend_plot)
```

```{r}
# Join the two feature plots
lime_linear_plots <- plot_grid(linear_plot_feature1, 
                               linear_plot_feature2, 
                               size_legend,
                               ncol = 3, 
                               rel_widths = c(0.4, 0.4, 0.1))
```

```{r}
# Create a plot to extract a legend for the prediction of interest
lime_legend_plot <- ggplot(data = prediction_of_interest, 
                           mapping = aes(x = feature1, 
                                         y = prediction, 
                                         fill = color)) +
  geom_point(size = 5, shape = 23) +
  scale_fill_manual(values = c("#EFB084")) +
  theme_classic(base_family = "Times") +
  labs(fill = "") +
  theme(legend.position = "bottom")

# Extract the legend
lime_legend <- get_legend(lime_legend_plot)
```

```{r}
# Join the title, plots, legend, and caption into one figure
lime_linear_concept <- plot_grid(lime_linear_plots, lime_legend,
                         ncol = 1,
                         rel_heights = c(0.9, 0.1))
```

This example shows that the explainer model is doing a good job of capturing the local trend.

```{r}
# View the plot
lime_linear_concept
```

### Ridge Regression Explainer

```{r}
# Fits the explainer model (weighted ridge regression)
ridge_explainer <- glmnet(x = lime_data %>% 
                            select(feature1, feature2) %>%
                            as.matrix(), 
                          y = lime_data %>% pull(prediction),
                          weights = lime_data$distance,
                          family = "gaussian", 
                          alpha = 0, 
                          lambda = 1, 
                          standardize = FALSE)
```

```{r}
# Extract the coefficients from the explainer
ridge_b0 <- ridge_explainer$a0[[1]]
ridge_b1 <- ridge_explainer$beta[1,1]
ridge_b2 <- ridge_explainer$beta[2,1]
```

```{r}
# Predictions versus feature 1
ridge_plot_feature1 <- ggplot(data = lime_data,
                        mapping = aes(x = feature1, 
                                      y = prediction, 
                                      size = distance)) + 
  geom_point(color = "grey30",
             alpha = 0.75) + 
  geom_abline(intercept = ridge_b0 +
                ridge_b2*prediction_of_interest$feature2, 
              slope = ridge_b1, 
              size = 1) +
  geom_point(data = prediction_of_interest,
             mapping = aes(x = feature1, y = prediction),
             fill = "#EFB084",
             color = "black",
             size = 5, 
             shape = 23) +
  theme_linedraw(base_family = "Times") +
  labs(x = "Feature 1", 
       y = "Black-Box Prediction") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none")
```

```{r}
# Predictions versus feature 2
ridge_plot_feature2 <- ggplot(data = lime_data,
                        mapping = aes(x = feature2, 
                                      y = prediction, 
                                      size = distance)) + 
  geom_point(color = "grey30",
             alpha = 0.75) + 
  geom_abline(intercept = ridge_b0 + 
                ridge_b1*prediction_of_interest$feature1,
              slope = ridge_b2,
              size = 1) +
  geom_point(data = prediction_of_interest,
             mapping = aes(x = feature2, y = prediction),
             fill = "#EFB084",
             color = "black",
             size = 5, 
             shape = 23) +
  theme_linedraw(base_family = "Times") +
  labs(x = "Feature 2", 
       y = "Black-Box Prediction") + 
  theme(panel.grid.major = element_blank(), 
        panel.grid.minor = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position = "none")
```

```{r}
# Join the two feature plots
lime_ridge_plots <- plot_grid(ridge_plot_feature1, 
                              ridge_plot_feature2, 
                              size_legend,
                              ncol = 3,
                              rel_widths = c(0.4, 0.4, 0.1))
```

```{r}
# Join the title, plots, legend, and caption into one figure
lime_ridge_concept <- plot_grid(lime_ridge_plots, 
                                lime_legend,
                                ncol = 1,
                                rel_heights = c(0.9, 0.1))
```

This example shows that the explainer model is **not** doing a good job of capturing the local trend.

```{r}
# Preview the plot
lime_ridge_concept
```
